--- 
title: "BSGatlas: Genetrack"
author:
  - name: Adrian Geissler
    affiliation:
    - &RTH Center for non-coding RNA in Technology and Health,
      Department of Veterinary and Animal Sciences, University of Copenhagen,
      Denmark
  - name: Christian Anthon
    affiliation: *RTH
  - name: Stefan Seemann
    affiliation: *RTH
  - name: Jan Gorodkin
    affiliation:
    - *RTH
    - Corresponding author
date: "`r Sys.time()`"
output:
  BiocStyle::html_document:
    self_contained: TRUE
    toc_float: TRUE
    css: pretty_table.css
---


```{r setup, include = FALSE}
source('00_load.R')

load('01_refseq.rda')
```

# Gene track

```{r, eval = FALSE}


```

The newest refseq annotation is from `r refseq$date`,
which is based on a major annotation refinment
(Borriss R, et al. Microb Biotechnol. 2018).
We downloaded the annotation in the genbank format
and parsed it with the bioconductor package `genbankr`.
We used simple string matching, to infer more specific 
non-coding RNA types. For only
`r refseq$noncoding %>% filter(type == 'unclear') %>% nrow`
of the
`r nrow(refseq$noncoding)` ncRNA genes it is not directly
clear from the annotaiton itself what they are.
(Maybe these can be refined in the merging)


SubtiWiki downloaded on June 3, 2019
BsubCyc version 38 (release Aug 9, 2017)

