--- 
title: "BSGatlas: Genetrack"
author:
  - name: Adrian Geissler
    affiliation:
    - &RTH Center for non-coding RNA in Technology and Health,
      Department of Veterinary and Animal Sciences, University of Copenhagen,
      Denmark
  - name: Christian Anthon
    affiliation: *RTH
  - name: Stefan Seemann
    affiliation: *RTH
  - name: Jan Gorodkin
    affiliation:
    - *RTH
    - Corresponding author
date: "`r Sys.time()`"
output:
  BiocStyle::html_document:
    self_contained: TRUE
    toc_float: TRUE
    css: pretty_table.css
---


```{r setup, include = FALSE}
source('00_load.R')

load('01_refseq.rda')
```

# Gene track

```{r, eval = FALSE}


```

The newest refseq annotation is from `r refseq$date`,
which is based on a major annotation refinment
(Borriss R, et al. Microb Biotechnol. 2018).
We downloaded the annotation in the genbank format
and parsed it with the bioconductor package `genbankr`.
We used simple string matching, to infer more specific 
non-coding RNA types. For only
`r refseq$noncoding %>% filter(type == 'unclear') %>% nrow`
of the
`r nrow(refseq$noncoding)` ncRNA genes it is not directly
clear from the annotaiton itself what they are.
(Maybe these can be refined in the merging)


SubtiWiki downloaded on June 3, 2019
BsubCyc version 38 (release Aug 9, 2017)


The 3016 families of rfam 14.1 were used to create ncRNA and structure annotations
based on conserved stucutre modles, as found by cmsearch v1.1.1.
(TODO: Why a low search space)
The search was conducted at three sensetiviy levels.
The conservative level had a maximal E-value cut-off at 1e-6 and required a
match score of at least the gathering score, a number provided for each
family by rfam.
At the medium level, the score requirement was reduced to half of the gathering
score, and at the senseitive level the E-value requirement was
relaxed to 10.
Afterwards, over-lapping matches, possibly by different rfam family models,
were combined and assigned to the family with the highest score.
(@CA is that correct)



